---
title: "Week 7"
author: "Sandeep Joshi"
date: "July 22, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, message=FALSE,cache=FALSE}
library("tidyverse")
```

## 3.2.4 Exercises
1. Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg)
```
This code creates an empty plot.


2. How many rows are in mpg? How many columns?
There are `r nrow(mtcars)` rows and `r ncol(mtcars)` columns.

3. What does the drv variable describe? Read the help for ?mpg to find out.
```{r}
?mpg
```
drv variable is for drive. i.e. f = front-wheel drive, r = rear wheel drive, 4 = 4wd


4. Make a scatterplot of hwy vs cyl.
```{r}
ggplot(mpg, aes(x = hwy, y = cyl)) + geom_point()
```


5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r}
ggplot(mpg, aes(x = class, y = drv)) + geom_point()
```
As we can see there are very few data points and these are two categorical datas. They do not have much correlations intutively as well. So we don't get any value out of plotting these.


## 3.3.1 Exercises
1. What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), colour = "blue")
```
This fixes it. We were giving color in aes hence it was gettting read as a variable with value blue and plotted as a point in legends in other words it was being used as a third dim say z for plot of x,y,z


2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

```{r}
glimpse(mpg)
```
Categorical: model, trans, drv, fl, class
Continuous: displ, year, cyl, cty, hwy


3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = cty, size = cty)) + geom_point()
ggplot(mpg, aes(x = displ, y = hwy, colour = fl, size = fl)) + geom_point()
```
As we can see from two plots if we use continuous variable we get same color with different shades/ intesity but when we use categorical data we get different colors for different values as there's no logical relation of shade/intensity or size with different values so we end up showing different colors or shapes. 


4. What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = hwy, size = hwy)) + geom_point()
```

There's not much value in denoting same data using two or more means. It might hamper readability and might lead to confusion.

5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}
?geom_point
# For shapes that have a border (like 21), you can colour the inside and
# outside separately. Use the stroke aesthetic to modify the width of the
# border
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 21, colour = "black", fill = "white", size = 5, stroke = 5)
ggplot(mtcars, aes(wt, mpg)) +
  geom_point(shape = 22, colour = "black", fill = "white", size = 5, stroke = 10)
```
stroke changes the brush stroke used to make border of the shapes. It only works with shapes with borders e.g 21, 22 etc


6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.
```{r}
ggplot(mpg, aes(x = displ, y = hwy, colour = displ < 5)) + geom_point()
```
We can see based on displ being greater than or less than 5 we have values TRUE or FALSE for color and R uses different colors to show these states.


## 3.5.1 Exercises
1. What happens if you facet on a continuous variable?
Things get messy. The continuous variable is converted to a categorical variable and R plots a facet for each distinct value.

2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv ~ cyl)
```
Empty cell means that there are no values for this combination of drv and cyl.

3. What plots does the following code make? What does . do?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```
The symbol `.` ignores that dimension when faceting. We can see in both graphs only faceting is done only on one dimension.
 
4. Take the first faceted plot in this section:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
Advantages of faceting over colour is that former is more readable in plots as less manglinging due to large datapoints or overlaps are there. Hard to discern between certains colors specially if you have color-blindness.

Disadvantages would be scaling down of graph due to multiple plots. Overlaying of plots makes comparisions easier.


5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?
```{r}
?facet_wrap
```
The arguments nrow determines the number of rows to use when using the facets.
It is necessary since facet_wrap() only facets on one variable.

The nrow as well as ncol arguments are unnecessary for facet_grid since the number of unique values of the variables specified in the function determines the number of rows and columns.



6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?
We have more space horizontally i.e. x axis in generic display (hardware and settings). If we put variables iwth more unique levels in columns they will be plotted in x axis so there'll be more space to denote and label them out.




## 3.6.1 Exercises
1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
line chart: geom_line
boxplot: geom_boxplot
histogram: geom_histogram
Area chart: geom_area

2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
Prediction: This code produces a scatter plot with displ on x-axis and hwy on  y-axis, with different points colored by different categorical values for drv. These will be fitted via smooth line without standard errors.

Conclusion: Ta-Da as predicted. 

3. What does show.legend = FALSE do? What happens if you remove it?
It hides the legend box. Nothing drastic happens but our readability of graph hampers. We may be not able to read what color denotes what line. Say in graph above we see colored line but we won't be able to tell which color correponds to which drv.

Why do you think I used it earlier in the chapter?
I think this usage refers to section 3.6 where we have three plots. We are only using aes color in last plot. Having legend there would have rescaled the plot and then we wouldn't find it as easy to visually compare the three plots.

4. What does the se argument to geom_smooth() do?
It adds standard error bands to the fit line.

5. Will these two graphs look different? Why/why not?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```
They will look same as we have exact same params for geom_point and geom_smooth in seconds as we had for ggplot in first. These arguments are carried over from ggplot and overridden in geom_point and geom_smooth. Since they have nothing, these end up being exactly same.

6. Recreate the R code necessary to generate the following graphs.
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(group = drv), se = FALSE) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy, colour = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(aes(linetype = drv), se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(size = 5, color = "white") +
  geom_point(aes(colour = drv))
```




## 3.7.1 Exercises
1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?
Default geom associated with stat_summary is geom_pointrange
```{r}
ggplot(data = diamonds) +
  geom_pointrange(mapping = aes(x = cut, y = depth), stat = "summary")
```

2. What does geom_col() do? How is it different to geom_bar()?
geom_col takes x and y data and sizes bars of x based on value of y. geom_bar just takes x and makes bar height based on counts of x for that value.
Also, the default stat of geom_col is stat_identity whereas default stat of geom_bar is stat_bin.


3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?
They have same trailing names post underscore in most of the cases(?). Refer: https://ggplot2.tidyverse.org/reference/#section-layer-stats, http://sape.inf.usi.ch/quick-reference/ggplot2/geom

4. What variables does stat_smooth() compute? What parameters control its behaviour?
```{r}
?stat_smooth
```

Computed variables

y: predicted value
ymin: lower pointwise confidence interval around the mean
ymax: upper pointwise confidence interval around the mean
se: standard error

Parameters which control its behaviour
mapping, method, formula, etcß


5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop..))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))
```
As seen when `group = 1` is not included, then all the bars in the plot have the same height, a height of 1. This is because geom_bar assumes that the groups are equal to the `x` values so each is its own group.

## 3.8.1 Exercises
1. What is the problem with this plot? How could you improve it?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```
There's overplotting going on in here so to avoid this gridding we can use jitter 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy), position = "jitter")
```



2. What parameters to geom_jitter() control the amount of jittering?
*width* controls the amount of vertical displacement
*height* controls the amount of horizontal displacement.

3. Compare and contrast geom_jitter() with geom_count().
geom_jitter adds noise to avoid overgridding and it changes the value of x and y slightly so it could be visually discernable and no overlapping occurs. 
geom_count on the other hand increase the size of the plot-point based on all the plot points that coincide there.

4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.
The default position adjustment for geom_boxplot is position_dodge2 or just dodge2.
```{r}
ggplot(data = mpg, aes(x = drv, y = hwy, colour = class)) +
  geom_boxplot()
```


## 3.9.1 Exercises
1. Turn a stacked bar chart into a pie chart using coord_polar().
```{r}
ggplot(mpg, aes(x = factor(1), fill = drv)) + geom_bar(width = 1) + coord_polar(theta = "y")
```
2. What does labs() do? Read the documentation.
```{r}
?labs
```
labs is for labels.
The labs function adds/ modifies axis titles, plot titles, and a caption to the plot.

3. What’s the difference between coord_quickmap() and coord_map()?
These both are used to project 3d earth to 2d plane. coord_quickmap is more approximate and error prone and is more viable for smaller areas which are close to equator as distoration in that area and scale is low. Compared to coord_map this is faster and less processing intense.


4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```
The function coord_fixed ensures that the line produced by geom_abline is at a 45-degree angle.
This adds a fixed reference line with slope 1 so we can visually compare the correlation.

