---
title: "Week 6"
author: "Sandeep Joshi"
date: "July 22, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---



## Exercise 5.2.1
1. Find all flights that

  1. Had an arrival delay of two or more hours
```{r}
library("nycflights13")
filter(flights, arr_delay >= 120)
```
  2. Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```
  3. Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```
  4. Departed in summer (July, August, and September)
```{r}
filter(flights, month %in% 7:9)
```
  5. Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```
  6. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```
  7. Departed between midnight and 6am (inclusive)
  We need to include midnight which is 2400 hrs. Other than that 0001 to 0600 hours are covered in <=
```{r}
filter(flights, dep_time <= 600 | dep_time == 2400)
```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

  1. Had an arrival delay of two or more hours
```{r}
filter(flights, between(arr_delay, 120, Inf))
```
 
  4. Departed in summer (July, August, and September)
```{r}
filter(flights, between(month, 7, 9))
```
  5. Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, between(arr_delay, 120, Inf), between(dep_delay, -Inf, 0))
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
filter(flights, is.na(dep_time))
```
Many such flights have arr_time also missing, these might be cancelled flights or off the logs flights
(tin-foil hats on time).

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
NA ^ 0 is universally 1 across all number range. So x or NA doesn't matter.
NA | TRUE is always TRUE as this is an OR statement and we only need one TRUE to make it so.
FALSE & NA is not missing as one ANDed FALSE would always result in FALSE.

##5.3.1 Exercises

1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```
In here, is.na would be 1 for missing values and 0 for non-missing ones. So on ordering that by desc we
will have NAs at top.

2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
t <- arrange(flights, desc(dep_delay))
```
Most delayed flight was `r t[1,]$carrier`'s `r t[1,]$flight` with tail-number `r t[1,]$tailnum`.
Flight that left earliers was `r t[nrow(t),]$carrier`'s `r t[nrow(t),]$flight` with tail-number `r t[nrow(t),]$tailnum`.

3. Sort flights to find the fastest flights.
```{r}
top_speed <- flights %>% mutate(mph = distance * 60 / air_time) %>% arrange(desc(mph))
```
Fastest flight was  `r top_speed[1,]$carrier`'s `r top_speed[1,]$flight` with tail-number `r top_speed[1,]$tailnum`.


4. Which flights travelled the longest? Which travelled the shortest?
```{r}
t <- arrange(flights, desc(distance))
```
Longest flight was `r t[1,]$carrier`'s `r t[1,]$flight` with tail-number `r t[1,]$tailnum`.
Shortest flith was `r t[nrow(t),]$carrier`'s `r t[nrow(t),]$flight` with tail-number `r t[nrow(t),]$tailnum`.


## 5.4.1 Exercises
Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
# Giving good old column names
```{r}
    select(flights, dep_time, dep_delay, arr_time, arr_delay)
```

# Giving column indices
```{r}
select(flights, 4, 6, 7, 9)
```

# Regex match which some what works in this case
```{r}
select(flights, starts_with("dep_"), starts_with("arr_"))
select(flights, matches("^(dep|arr)_(time|delay)$"))
```

What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, dep_delay, dep_delay, dep_delay)
select(flights, 1, 2, 2, 1)

```
It seems that it ignores the duplication of variables

What does the one_of() function do? Why might it be helpful in conjunction with this vector?
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
**one_of** is useful to pass in quoted character vector of variable names instead of passing unquoted variable name arguments. This is useful in dynamic or progamatically passing variables say to a function.


Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?

select(flights, contains("TIME"))
```{r}
select(flights, contains("TIME"))
```
**No it doesn't coz I haz the smarts!!**
Contains matches string (i.e TIME) with case-agnosticism so all variables (columns) with aNY CaSE of TiME in them are pulled up.

ignore.case is required to set to FALSE to change that default
```{r}
select(flights, contains("TIME", ignore.case = F))
```


## 5.5.2 Exercises
Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
flights_times <- flights %>% mutate(dep_time_mins = strtoi(as.difftime(dep_time, format = "%H:%M:%S", units = "mins")),
                                    sched_dep_time_mins = strtoi(as.difftime(sched_dep_time, format = "%H:%M:%S", units = "mins"))) %>% select(dep_time, dep_time_mins, sched_dep_time, sched_dep_time_mins)
glimpse(flights_times)
head(flights_times)
```


Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
Mathematically, air_time = arr_time - dep_time
```{r}
airtime <- flights %>% mutate(air_time_diff = air_time - arr_time + dep_time)
nrow(filter(airtime, air_time_diff != 0))
nrow(filter(airtime, air_time_diff < 0))
nrow(filter(airtime, air_time_diff > 0))
```

We can see that these are not equal. air_time might not include taxing time for plane which is included in arr_time and dep_times.
We are not considering the fact when a day has elapsed and computation for arr_time - dep_time would not even be positive in many of these cases! Which is totally incorrect.

We can fix this by using more sophisticated way of find arr_time - dep_time and working with timestamps and Posix or epoch and not just military hours without the sense of day/ date.

Also, we are not even sure if the times stated are in single timezone or not.


Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
I expect them to follow this equation:
*dep_time - sched_dep_time = dep_delay*

Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().
```{r}
flights_delayed <- flights %>% mutate(dep_delay_min_rank = min_rank(desc(dep_delay))) %>% arrange(dep_delay_min_rank)
head(flights_delayed, 10)
```


What does 1:3 + 1:10 return? Why?
```{r}
1:3 + 1:10
```
It gives a warning as shorter vector on being repeated 3 only gives us 9 elements whereas larger vector
has 10, so it doesn't completly cover all elements of larger vector.
so what we have effectively is: c(rep(1:3,3),1) + 1:10

What trigonometric functions does R provide?
```{r}
?Trig
```
sin, cos, tan, acos, asin, atan, atan1, cospi, sinpi and tanpi. These respectively compute the sine, cosine, tangent, arc-cosine, arc-sine, arc-tangent, and the two-argument arc-tangent.




## 5.6.7 Exercises
1 Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

  A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.

  A flight is always 10 minutes late.

  A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

  99% of the time a flight is on time. 1% of the time it’s 2 hours late.

  Which is more important: arrival delay or departure delay?
  
I'd want to compare the minutes lost for all passangers for which we would need some more flight data and would try to make a function which minimizes this value. Say, flight A carrying 2 passanger delayed by 120 minutes (240 minutes total) may still be better than flight B delayed by 10 minutes carrying 100 passengers (1000 minutes total).

We can add some value to delay function which might rise by some quantified margins after a bit, as in after delay of 2 hours we might need to accomodate passengers with missed connections and/or refund vs minor delay of 20 minutes 50% of time.


2 Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).
```{r}
not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% count(dest)
not_cancelled %>% count(tailnum, wt = distance) 

not_cancelled %>% group_by(dest) %>% tally()
not_cancelled %>% group_by(tailnum) %>% summarise(n = sum(distance))
```


3 Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?
There might be other ways we might not have this data. A cancelled flight would not have taken off so we can just look at the departure time instead. is.na(dep_time)

4 Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?
```{r}
cancelled_per_day <- flights %>% mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>% group_by(year, month, day) %>% summarise(mean_cancelled = mean(cancelled), 
                                         mean_dep_delay = mean(dep_delay, na.rm = TRUE),
                                         mean_arr_delay = mean(arr_delay, na.rm = TRUE)
                                         ) 
cancelled_per_day
```

```{r}
ggplot(cancelled_per_day) + geom_point(aes(x = mean_dep_delay, y = mean_cancelled))
```

```{r}
ggplot(cancelled_per_day) + geom_point(aes(x = mean_arr_delay, y = mean_cancelled))
```

Both these delays show a upward increasing relation to cancellation. So more chances of cancellation on increase of any of these delays.

5 Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

Comparing on arrival delays 
```{r}
flights %>% group_by(carrier, dest) %>% summarise(arr_delay = mean(arr_delay, na.rm = TRUE)) %>% arrange(desc(arr_delay))

flights %>% group_by(carrier, dest) %>% summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>% arrange(desc(dep_delay))
```
We do see some repeat offenders and bad combinations here but it seems that UA and OO are typically bad across different Airports.


6 What does the sort argument to count() do. When might you use it?
Sort argument, if TRUE will sort output in descending order of n



## 5.7.1 Exercises
1. Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.
Mostly summary/ aggregation functions like Mean, Avg, max, min etc are affected by grouping. When using grouping these are performed in a group and not across the whole dataset.

2. Which plane (tailnum) has the worst on-time record?
I'll compare the dep delays to get this. So getting max average dep delay across tailnums
```{r}
flights %>%
  group_by(tailnum) %>%
  summarise(mean_dep_delay = mean(dep_delay), n = n()) %>% filter(min_rank(desc(mean_dep_delay)) == 1)
```
Here we only see one instance of it being late, if we have filter for n atleast be 10 or more, we get a different flight.
```{r}
flights %>%
  group_by(tailnum) %>%
  summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE), n = n()) %>% filter(n >= 10) %>% filter(min_rank(desc(mean_dep_delay)) == 1)
```

3. What time of day should you fly if you want to avoid delays as much as possible?
```{r}
flights %>% group_by(hour) %>% summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE)) %>% arrange(mean_dep_delay)
```
So 5 in the AM sounds amazing!

4. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.
```{r}
flights %>%
  filter(arr_delay > 0) %>%
  group_by(dest) %>%
  mutate(
    arr_delay_total = sum(arr_delay),
    arr_delay_prop = arr_delay / arr_delay_total
  ) %>%
  select(dest, month, day, dep_time, carrier, flight,
         arr_delay, arr_delay_prop) %>%
  arrange(dest, desc(arr_delay_prop))
```

5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag(), explore how the delay of a flight is related to the delay of the immediately preceding flight.
```{r}
lagged_delays <- flights %>%
  arrange(origin, month, day, dep_time) %>%
  group_by(origin) %>%
  mutate(dep_delay_lag = lag(dep_delay)) %>%
  filter(!is.na(dep_delay), !is.na(dep_delay_lag))
```

Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?


7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.
# Ranking carriers based on? I am choosing the destinations they fly to.
```{r}
flights %>% 
  # Gives us destinations with more than 1 carriers
  group_by(dest) %>% mutate(no_carriers = n_distinct(carrier)) %>% filter(no_carriers > 1) %>%
   # Destinations served by that carrier
   group_by(carrier) %>% summarise(no_dest = n_distinct(dest)) %>% arrange(desc(no_dest))
```

8. For each plane, count the number of flights before the first delay of greater than 1 hour.
```{r}
flights %>%
  # sort in increasing order
  filter(!is.na(dep_delay)) %>% arrange(tailnum, year, month, day) %>% group_by(tailnum) %>%
  # cumulative number of flights delayed over one hour
  mutate(cumulative_delays = cumsum(dep_delay > 60)) %>%
  # count the number of flights == 0
  summarise(total_flights = sum(cumulative_delays < 1)) %>% arrange(total_flights)
```
