---
title: "Week 11"
author: "Sandeep Joshi"
date: "July 22, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compute the Principal Components
```{r}
mtcars.pca <- prcomp(mtcars[,c(1:7,10,11)], center = TRUE,scale. = TRUE)
summary(mtcars.pca)
str(mtcars.pca)
```

## Plotting PCA
```{r}
# Install required package of ggbiplot
install.packages("devtools", repos = "http://cran.us.r-project.org")
library(devtools)
install_github("vqv/ggbiplot")

library(ggbiplot)
# Plot the PCA
ggbiplot(mtcars.pca)

# Plot again with labels for cars
ggbiplot(mtcars.pca, labels=rownames(mtcars))
```

## Interpreting the results
```{r}
# Categorizing cars on their origins
mtcars.country <- c(rep("Japan", 3), rep("US",4), rep("Europe", 7),
                    rep("US",3), "Europe", rep("Japan", 3), rep("US",4), 
                    rep("Europe", 3), "US", rep("Europe", 3))
ggbiplot(mtcars.pca,ellipse=TRUE,  labels=rownames(mtcars), groups=mtcars.country)

# PC3 and PC4
ggbiplot(mtcars.pca,ellipse=TRUE,choices=c(3,4),   labels=rownames(mtcars), groups=mtcars.country)
```

## Playing with ggbiplot features
```{r}
# Using circle argument for the centre of the dataset
ggbiplot(mtcars.pca,ellipse=TRUE,circle=TRUE, labels=rownames(mtcars), groups=mtcars.country)

# Using custom scaling for observations and variables
ggbiplot(mtcars.pca,ellipse=TRUE,obs.scale = 1, var.scale = 1,  labels=rownames(mtcars), groups=mtcars.country)

# Removing the arrows using var.axes
ggbiplot(mtcars.pca,ellipse=TRUE,obs.scale = 1, var.scale = 1,var.axes=FALSE,   labels=rownames(mtcars), groups=mtcars.country)

# Custom ggbiplot with: custom colors for groups, title, in built theme and repositioning legend.
ggbiplot(mtcars.pca,ellipse=TRUE,obs.scale = 1, var.scale = 1,  labels=rownames(mtcars), groups=mtcars.country) +
  scale_colour_manual(name="Origin", values= c("forest green", "red3", "dark blue"))+
  ggtitle("PCA of mtcars dataset")+
  theme_minimal()+
  theme(legend.position = "bottom")
```

## Adding another unearthly sample to the dataset
```{r}
spacecar <- c(1000,60,50,500,0,0.5,2.5,0,1,0,0)

mtcarsplus <- rbind(mtcars, spacecar)
mtcars.countryplus <- c(mtcars.country, "Jupiter")

mtcarsplus.pca <- prcomp(mtcarsplus[,c(1:7,10,11)], center = TRUE,scale. = TRUE)

ggbiplot(mtcarsplus.pca, obs.scale = 1, var.scale = 1, ellipse = TRUE, circle = FALSE, var.axes=TRUE, labels=c(rownames(mtcars), "spacecar"), groups=mtcars.countryplus)+
  scale_colour_manual(name="Origin", values= c("forest green", "red3", "violet", "dark blue"))+
  ggtitle("PCA of mtcars dataset, with extra sample added")+
  theme_minimal()+
  theme(legend.position = "bottom")
```
# We can see that the unearthly specs of new spacecar has skewed our PCA a lot.

## Projecting this new sample to original PCA
```{r}
s.sc <- scale(t(spacecar[c(1:7,10,11)]), center= mtcars.pca$center)
s.pred <- s.sc %*% mtcars.pca$rotation


mtcars.plusproj.pca <- mtcars.pca
mtcars.plusproj.pca$x <- rbind(mtcars.plusproj.pca$x, s.pred)


ggbiplot(mtcars.plusproj.pca, obs.scale = 1, var.scale = 1, ellipse = TRUE, circle = FALSE, var.axes=TRUE, labels=c(rownames(mtcars), "spacecar"), groups=mtcars.countryplus)+
  scale_colour_manual(name="Origin", values= c("forest green", "red3", "violet", "dark blue"))+
  ggtitle("PCA of mtcars dataset, with extra sample projected")+
  theme_minimal()+
  theme(legend.position = "bottom")
```
# Now our PCA is still showing the new Spacecar but since we projected this to older PCA that was done # with out it, it no longer skews the other samples but also it cannot be assigned to any groups.









